---
title: "1. Biodiversity Index Variation - Start"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
This Rmarkdown notebook is intended to serve as a useful starting place for this Zambia Biodiversity Index Variation project. It was originally intended for development, but is not strictly necessary for running the primary analyses. Start with "2_BD_input_prep.Rmd."

This notebook includes chunks that:
- load packages, functions, pathnames, etc.
- load some basic, useful gis files (these are sometimes recreated or superceded in later .Rmd files)
- explore the inner workings of the `agroEcoTradeoff` model, specifically the way results are stored

```{r start-up, eval=TRUE}
## libraries --------------------------------------------------------------------------------
source("scripts/cc_libraries.R")
source("scripts/cc_functions.R")
source("scripts/cc_pathnames.R")
```

```{r load-basic-gis-files}

# Other helpful files ----------------------------------------------------------------
aaeac <- sf::st_crs("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs") # Africa Albers Equal Area Conic projection.
# could probably also use raster::crs
CRSobj <- sp::CRS("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") # set CRS object for model runs, noting that this must be a CRS object for sp*, not a crs object
class(aaeac)
class(CRSobj)

crs_longlat <- sf::st_crs("+proj=longlat +datum=WGS84 +no_defs")

load(file = fp(p_ZA,"parks_roads.rda")) # includes roads (a SpatialLinesDataFrame), pas (SpatialPolygonsDataFrame, a shapefile that includes both national parks and GMAs), and zambia (SpatialPolygonsDataFrame, outline of Zambia)
msk_shp <- readOGR(fp(p_datnew,"msk.shp")) %>% 
  spTransform(CRS("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
msk_sf <- st_read(fp(p_datnew,"msk.shp")) %>% st_transform(aaeac)
msk_sf_noholes <- smoothr::fill_holes(msk_sf, threshold = units::set_units(10000, km^2)) # a method to fill holes. could also use spatialEco::remove.holes() on sp* type data.
msk_sf_buff <- st_buffer(msk_sf, units::set_units(100, km))

msk_sf_ll <- st_transform(msk_sf, crs_longlat)
msk_sf_ll_noholes <- st_transform(msk_sf_noholes, crs_longlat)
msk_sf_ll_buff <- st_buffer(msk_sf_ll, units::set_units(1, degree))

msk <- raster(fp(p_datnew,"msk.tif")) # the actual high resolution msk, created directly from the csv in "bd_new_prep.Rmd"
# also equal to this: plot(dt_to_raster(il$mask, CRSobj))

# files from Estes model run
cp <- read.csv(fp(p_ZA,"ZA-cons-priorities.csv"))
cp.dt <- fread(fp(p_ZA,"ZA-cons-priorities.csv"))

cp_raster <- raster(fp(p_datnew,"cons-priorities_from_csv.tif"))
```

```{r inputs_agroEcoTradeoff}
# -----------------------------------------------------------------------
# inputs forand from agroEcoTradeoff model
# -----------------------------------------------------------------------
il <- agroEcoTradeoff::fetch_inputs(path = p_ZA)
tradeoff_batch()
tradeoff_mod()
    # conv <- constraints(inlist = list(Y = il$y_std, C = il$carbon_p, 
    #     BD = il$cons_p, COST = il$cost_p), cbetas = cbetas, silent = silent)
    # converted <- convert(conv_prob = conv, target = target, pot_yield = il$p_yield, 
    #     cropnames = il$cropnames, base = il$mask, ha = il$sp$ha, 
    #     keep_index = FALSE)
# standardize()


msk_csv <- il$mask[,-"ind"] # msk_csv, for use in creating rasters later on
## note: this is identical to il$mask
# msk_csv <- fread(paste0(p_dat, .Platform$file.sep,
#                         "ZA", .Platform$file.sep,
#                         "ZA", "-mask.csv"))

agroEcoTradeoff::input_handler()
agroEcoTradeoff::standardize()

# load convertible areas - this is a layer of the proportion of each cell that is not in ag or urban land use (i.e. the proportion of land that is convertible).  This is the same as the file listed as an input in the tradeoff model outputs, e.g. "reference_y_toff$inputs$convertible"

#convertible <- fread(fp(p_dat, "ZA/ZA-convertible.csv"))

convertible_r <- il$convertible %>%
  cbind(msk_csv, .) %>%
  dt_to_raster(., CRSobj)

# convertible without the Protected Areas (PAs)

intpa <- fread(paste0(p_dat, .Platform$file.sep,
                        "ZA", .Platform$file.sep,
                        "ZA", "-intpa.csv"))

valinds <- which(il$convertible > 0)

il$convertible[valinds, ] # selecting only cells that are at least partially convertible (i.e. outside of pas and entirely urban areas). This is the data.table method for masking
il$convertible[which(il$convertible > 0), ]

# to make a new convertible raster, excluding all of the cells with a value of 0:
convertible <- cbind(il$mask, il$convertible)
convertible[valinds, ]
convertible_r0 <- dt_to_raster(convertible[valinds, ], CRSobj)$convertible

test <- il$convertible[valinds, ] %>% 
  cbind(msk_csv_no_pas) %>% 
  dt_to_raster(CRSobj)

plot(test$convertible)
plot(convertible_r0)
identical(test$convertible, convertible_r0)

# can also just manually set cells with 0 to NA:
test <- convertible_r
test[test == 0] <- NA
plot(test)


dt_m[,3]
msk_csv_no_pas <- msk_csv[valinds, ]
plot(dt_to_raster(msk_csv_no_pas, CRSobj))
plot(dt_to_raster(convertible, CRSobj)$convertible)


dt_m[, 3]
plot(facet_r_b$all_b)
# plot(convertible_r)

plot(dt_to_raster(cbind(msk_csv_no_pas, il$convertible[valinds, ]), CRSobj))
il$convertible[valinds, ] 

```


