---
title: "1. Biodiversity Index Variation - Start"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
This Rmarkdown notebook is intended to serve as a useful starting place for this Zambia Biodiversity Index Variation project. It includes chunks to:
- load packages, functions, pathnames, etc.
- load some basic, useful gis files (these are sometimes recreated or superceded in later .Rmd files)
- explore the inner workings of the `agroEcoTradeoff` model, specifically the way results are stored
- load prepared biodiversity input layers, including vertebrate species files, plants and habitats files, and final inputs and tradeoff model results
- model run naming conventions

```{r start-up, eval=TRUE}
## libraries --------------------------------------------------------------------------------
source("scripts/cc_libraries.R")
source("scripts/cc_functions.R")
source("scripts/cc_pathnames.R")
```

```{r load-basic-gis-files}

# Other helpful files ----------------------------------------------------------------
aaeac <- sf::st_crs("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs") # Africa Albers Equal Area Conic projection.
# could probably also use raster::crs
CRSobj <- sp::CRS("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") # set CRS object for model runs, noting that this must be a CRS object for sp*, not a crs object
class(aaeac)
class(CRSobj)

crs_longlat <- sf::st_crs("+proj=longlat +datum=WGS84 +no_defs")

load(file = fp(p_ZA,"parks_roads.rda")) # includes roads (a SpatialLinesDataFrame), pas (SpatialPolygonsDataFrame, a shapefile that includes both national parks and GMAs), and zambia (SpatialPolygonsDataFrame, outline of Zambia)
msk_shp <- readOGR(fp(p_datnew,"msk.shp")) %>% 
  spTransform(CRS("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
msk_sf <- st_read(fp(p_datnew,"msk.shp")) %>% st_transform(aaeac)
msk_sf_noholes <- smoothr::fill_holes(msk_sf, threshold = units::set_units(10000, km^2)) # a method to fill holes. could also use spatialEco::remove.holes() on sp* type data.
msk_sf_buff <- st_buffer(msk_sf, units::set_units(100, km))

msk_sf_ll <- st_transform(msk_sf, crs_longlat)
msk_sf_ll_noholes <- st_transform(msk_sf_noholes, crs_longlat)
msk_sf_ll_buff <- st_buffer(msk_sf_ll, units::set_units(1, degree))

msk <- raster(fp(p_datnew,"msk.tif")) # the actual high resolution msk, created directly from the csv in "bd_new_prep.Rmd"
# also equal to this: plot(dt_to_raster(il$mask, CRSobj))

# files from Estes model run
cp <- read.csv(fp(p_ZA,"ZA-cons-priorities.csv"))
cp.dt <- fread(fp(p_ZA,"ZA-cons-priorities.csv"))

cp_raster <- raster(fp(p_datnew,"cons-priorities_from_csv.tif"))
```

```{r inputs_agroEcoTradeoff}
# -----------------------------------------------------------------------
# inputs forand from agroEcoTradeoff model
# -----------------------------------------------------------------------
il <- agroEcoTradeoff::fetch_inputs(path = p_ZA)
tradeoff_batch()
tradeoff_mod()
    # conv <- constraints(inlist = list(Y = il$y_std, C = il$carbon_p, 
    #     BD = il$cons_p, COST = il$cost_p), cbetas = cbetas, silent = silent)
    # converted <- convert(conv_prob = conv, target = target, pot_yield = il$p_yield, 
    #     cropnames = il$cropnames, base = il$mask, ha = il$sp$ha, 
    #     keep_index = FALSE)
# standardize()


msk_csv <- il$mask[,-"ind"] # msk_csv, for use in creating rasters later on
## note: this is identical to il$mask
# msk_csv <- fread(paste0(p_dat, .Platform$file.sep,
#                         "ZA", .Platform$file.sep,
#                         "ZA", "-mask.csv"))

agroEcoTradeoff::input_handler()
agroEcoTradeoff::standardize()

# load convertible areas - this is a layer of the proportion of each cell that is not in ag or urban land use (i.e. the proportion of land that is convertible).  This is the same as the file listed as an input in the tradeoff model outputs, e.g. "reference_y_toff$inputs$convertible"

#convertible <- fread(fp(p_dat, "ZA/ZA-convertible.csv"))

convertible_r <- il$convertible %>%
  cbind(msk_csv, .) %>%
  dt_to_raster(., CRSobj)

# convertible without the Protected Areas (PAs)

intpa <- fread(paste0(p_dat, .Platform$file.sep,
                        "ZA", .Platform$file.sep,
                        "ZA", "-intpa.csv"))

valinds <- which(il$convertible > 0)

il$convertible[valinds, ] # selecting only cells that are at least partially convertible (i.e. outside of pas and entirely urban areas). This is the data.table method for masking
il$convertible[which(il$convertible > 0), ]

# to make a new convertible raster, excluding all of the cells with a value of 0:
convertible <- cbind(il$mask, il$convertible)
convertible[valinds, ]
convertible_r0 <- dt_to_raster(convertible[valinds, ], CRSobj)$convertible

test <- il$convertible[valinds, ] %>% 
  cbind(msk_csv_no_pas) %>% 
  dt_to_raster(CRSobj)

plot(test$convertible)
plot(convertible_r0)
identical(test$convertible, convertible_r0)

# can also just manually set cells with 0 to NA:
test <- convertible_r
test[test == 0] <- NA
plot(test)


dt_m[,3]
msk_csv_no_pas <- msk_csv[valinds, ]
plot(dt_to_raster(msk_csv_no_pas, CRSobj))
plot(dt_to_raster(convertible, CRSobj)$convertible)


dt_m[, 3]
plot(facet_r_b$all_b)
# plot(convertible_r)

plot(dt_to_raster(cbind(msk_csv_no_pas, il$convertible[valinds, ]), CRSobj))
il$convertible[valinds, ] 

```


```{r load_vert_files}
# final layers:
# load-vert-bricks ----------------------------------------------------------------------------
# create layer names object
layer_names <- c("mam", "bird", "amp", "rep", "sum", "sum_norm", "norm_sum", "mam_10", "bird_10", "amp_10", "rep_10", "norm_sum_10", "mam_110", "bird_110", "amp_110", "rep_110", "norm_sum_110")

richness_names <- c("all_richness", "endemism_richness", "endemism_zam_richness", "threat_richness", "threat_weighted_richness", "small_richness", "small_zam_richness", "small_threat_richness")

# load in as a list of all rasters
vert_r <- list(
  all_richness = brick(paste0(p_iucn_dev, "/", richness_names[1], ".tif")),
  endemism_richness = brick(paste0(p_iucn_dev, "/", richness_names[2], ".tif")),
  endemism_zam_richness = brick(paste0(p_iucn_dev, "/", richness_names[3], ".tif")),
  threat_richness = brick(paste0(p_iucn_dev, "/", richness_names[4], ".tif")),
  threat_weighted_richness = brick(paste0(p_iucn_dev, "/", richness_names[5], ".tif")),
  small_richness = brick(paste0(p_iucn_dev, "/", richness_names[6], ".tif")),
  small_zam_richness = brick(paste0(p_iucn_dev, "/", richness_names[7], ".tif")),
  small_threat_richness = brick(paste0(p_iucn_dev, "/", richness_names[8], ".tif"))
  )

# rename layers
for(i in seq_along(vert_r)) {
  names(vert_r[[i]]) <- layer_names
}


# vertebrate files -------------------------------------------------------------------------------------
# valid sf files, made valid, and prepped for use in small ranged species analyses
load(file = fp(p_iucn_dev,"mam_valid_prepped.RData"), verbose = TRUE)
load(file = fp(p_iucn_dev,"bird_valid_prepped.RData"), verbose = TRUE)
load(file = fp(p_iucn_dev,"amp_valid_prepped.RData"), verbose = TRUE)
load(file = fp(p_iucn_dev,"rep_valid_prepped.RData"), verbose = TRUE)

# vert zambia sf files, prepped and valid
load(file = fp(p_iucn_dev, "mam_zambia.RData"), verbose = TRUE) # includes mam_zambia, which has had area added, been prepped, and then clipped to zambia
load(file = fp(p_iucn_dev, "bird_zambia.RData"), verbose = TRUE) # includes bird_zambia, which has had area added, been prepped, and then clipped to zambia
load(file = fp(p_iucn_dev, "amp_zambia.RData"), verbose = TRUE) # includes amp_zambia, which has had area added, been prepped, and then clipped to zambia
load(file = fp(p_iucn_dev, "rep_zambia.RData"), verbose = TRUE) # includes rep_zambia, which has had area added, been prepped, and then clipped to zambia


# vertebrate lists - important -----------------
load(file = fp(p_iucn_dev, "vert_lists.RData"), verbose = TRUE)

load(file = fp(p_iucn_dev, "mam_list.RData"))
load(file = fp(p_iucn_dev, "bird_list.RData"))
load(file = fp(p_iucn_dev, "amp_list.RData"))
load(file = fp(p_iucn_dev, "rep_list.RData")) # note, this rep_list file includes only data from IUCN, not from GARD. See GARD list for the most up to date files. The richness metric .tifs have all been updated to use GARD (or IUCN in the case of threatened species richness). These are the finaly data files used in the analysis.

load(file = "/Users/christophercrawford/Google Drive/_Projects/Zambia/agroEcoTradeoff/external/data_new/1_IUCN_dev/gard_list.RData", verbose = TRUE) #



# load-vert-bricks -------------------------------------------------------------------------------
load(file = fp(p_datnew, "vert_bricks.Rdata"), verbose = TRUE) # this includes mam_brick, bird_brick, amp_brick, rep_brick
```


```{r load-plants-habitats}
# plants ------------------------------------------------------------------------
plants_zambia_r <- raster(fp(p_plants_dev, "plants_zambia_r.tif"))


# habitats rasters -----------------------------------------------------------------------------
load(file = fp(p_datnew,"habitat_rasters_zambia.RData"), verbose = TRUE)

ecoregions_weighted_rarity <- raster(fp(p_ecoreg_dev, "ecoregions_weighted_rarity.tif"))

# final zambia rasters:
load(file = fp(p_datnew,"final-rasters-zambia.RData"), verbose = TRUE)
```

```{r load-inputs+toffs}
load(file = fp(p_mod_inputs,"bd_inputs.RData"), verbose = TRUE)

bd_inputs_brick <- brick(fp(p_mod_inputs,"bd_inputs_brick.tif")); names(bd_inputs_brick) <- runs
# load(file = fp(p_mod_inputs,"bd_inputs.RData"), verbose = TRUE) # or, load individually
bd_inputs_brick


load(file = fp(p_mod_output, "reference_files.RData"), verbose = TRUE) # contains: reference_y_toff, reference_y_toff_outputs, reference_yct_toff, reference_yct_toff_outputs, reference_y, reference_yct,


bd_inputs_brick <- brick(fp(p_mod_inputs,"bd_inputs_brick.tif"))
names(bd_inputs_brick) <- runs

```


```{r *runs}
# Final model runs:

# 38 runs x 3 resolutions = 114 runs

runs_1 <- c(
  "vert_all",  # mam, bird, amp, and rep (comprehensive)
  "vert_endemism", # mam, bird, amp, and rep (comprehensive)
  "vert_threat", # mam, bird, amp, and rep (but only those reps that were assessed)
  "vert_small", # mam, bird, amp, and rep (comprehensive)
  "mam_all", 
  "bird_all", 
  "amp_all", 
  "rep_all",  # comprehensive
  "plants", 
  "mam_endemism",
  "bird_endemism", 
  "amp_endemism",  
  "rep_endemism",   # comprehensive
  "mam_threat",  
  "bird_threat", 
  "amp_threat", 
  "rep_threat",  # non comprehensive
  "mam_small", 
  "bird_small", 
  "amp_small", 
  "rep_small",  # comprehensive
  "average_mb",  
  "average_vp",  
  "average_ae",
  "geometric_mb", 
  "geometric_vp",
  "geometric_ae", 
  "max_mb", 
  "max_vp", 
  "max_ae",
  "multi_mb", 
  "multi_vp", 
  "multi_ae", 
  "estes", 
  "laurance", 
  "habitats",
  "bird_composite",
  "damania"
)
runs_1
runs_10 <- paste0(runs_1, "_10")
runs_110 <- paste0(runs_1, "_110")

# combine runs at the three resolutions
runs <- c(runs_1, runs_10, runs_110)


# extra runs, for reference:
extra_runs <- c("vert_all_mba", "vert_endemism_mba", "vert_threat_mba", "vert_small_mba", "vert_all_mba_10", "vert_endemism_mba_10", "vert_threat_mba_10", "vert_small_mba_10", "vert_all_mba_110", "vert_endemism_mba_110", "vert_threat_mba_110", "vert_small_mba_110", "vert_all_sum_norm", "vert_endemism_sum_norm", "vert_small_sum_norm", "vert_all_sum_norm_10", "vert_endemism_sum_norm_10", "vert_small_sum_norm_10", "vert_all_sum_norm_110", "vert_endemism_sum_norm_110", "vert_small_sum_norm_110")



brick_mod_names <- c(runs, "reference_yct", "reference_y")
runs_w_ref <- c(runs, "reference_yct", "reference_y")
runs_w_raw <- c(runs, paste0(runs[1:8], "_raw"))
```


```{r ensembles-names}
# ensemble names
ensemble_names <- c(
  "en1_types",
  "en1_types_b",
  
  "en2_taxa_all",
  "en2_taxa_endemism",
  "en2_taxa_threat",
  "en2_taxa_small",
  
  "en3_mb",
  "en3_vp",
  "en3_ae",
  
  "en4_all",
  "en4_endemism",
  "en4_threat",
  "en4_small",
  
  "en5_composites",
  
  "en6_norm_all",
  "en6_norm_endemism",
  "en6_norm_threat",
  "en6_norm_small",
  
  "all")

runs_w_ensembles <- c(runs_w_ref, ensemble_names)

bd_dt_r_names <- c(runs_w_raw, ensemble_names)

runs_w_ref_order <- as.factor(runs_w_ref)

levels(runs_w_ref_order)


# runs_reorder <- c(1, 21, 25, 2, 22, 26, 3, 23, 27, 4, 24, 28, 5:13, 15, 17, 19, 16, 18, 20, 14, 29, 30, 31, 32, 33, 34, 35) # ordering
runs
runs_reorder <- c(1:12, 14:48, 50, 13, 49) # ordering
runs_reorder <- c(46, 47, 50, 48, 49, 2, 13, 1, 3:12, 14:45, 51:54)
runs_reorder_plot <- c(1:4, 18:21, 26:29, 3, 40, 44, 46, 47, 50, 48, 49, 2, 13
                       #, 5:12, 14:17, 22:25, 30:39, 41:43, 45, 51:54
                       )

runs[runs_reorder_plot]

runs[en1_taxa_threat]

# -----------------------------------------------------------
# ensemble_names

en1_types <- c(1:4)
en1_types_b <- c(5:8)

en2_taxa_all <- c(9:13)
en2_taxa_endemism <- c(14:17)
en2_taxa_threat <- c(18:21)
en2_taxa_small <- c(22:25)

en3_mb <- c(26, 29, 32, 35)
en3_vp <- c(27, 30, 33, 36)
en3_ae <- c(28, 31, 34, 37)

en4_all <- c(1, 38, 42)
en4_endemism <- c(2, 39, 43)
en4_threat <- c(3, 40, 44)
en4_small <- c(4, 41, 45)

runs[ensembles_list[[14]]]

en5_composites <- c(46, 47, 50, 48, 49, 2, 13) # including vert_endemism and plants

en6_norm_all <- c(1, 51)
en6_norm_endemism <- c(2, 52)
en6_norm_threat <- c(3, 53)
en6_norm_small <- c(4, 54)

ensembles_list <- list(
  en1_types,
  en1_types_b,
  
  en2_taxa_all,
  en2_taxa_endemism,
  en2_taxa_threat,
  en2_taxa_small,
  
  en3_mb,
  en3_vp,
  en3_ae,
  
  en4_all,
  en4_endemism,
  en4_threat,
  en4_small,
  
  en5_composites,
  
  en6_norm_all,
  en6_norm_endemism,
  en6_norm_threat,
  en6_norm_small,
  
  seq_along(runs)
)


names(ensembles_list) <- ensemble_names




ensemble_col <- c(rep("en1", 12), rep("en4", 2), rep("en2", 6), rep("en3", 8), rep("en4", 3))


runs[en1_types]
runs[en1_types_b]

runs[en2_taxa_all]
runs[en2_taxa_endemism]
runs[en2_taxa_threat]
runs[en2_taxa_small]

runs[en3_mb]
runs[en3_vp]
runs[en3_ae]

runs[en4_all]
runs[ en4_endemism]
runs[en4_threat]
runs[en4_small]

runs[en5_composites]

runs[en6_norm_all]
runs[en6_norm_endemism]
runs[en6_norm_threat]
runs[en6_norm_small]


# ---------------------------------------------------------------------------------------------------------
# new ensembles, for reference:

# 54 layers

en1_types <- c(1:4)
# 
# 1.	vert_all
# 2.	vert_endemism
# 3.	vert_threat
# 4.	vert_small
# 
en1_types_b <- c(5:8)
# 
# 5.	vert_threat_weighted
# 6.	vert_small_threat
# 7.	vert_small_zam
# 8.	vert_endemism_zam
# 
en2_taxa_all <- c(9:13)
# 9.	mam_all
# 10.	bird_all
# 11.	amp_all
# 12.	rep_all
# 13.	plants
# 
en2_taxa_endemism <- c(14:17)
# 14.	mam_endemism
# 15.	bird_endemism
# 16.	amp_endemism
# 17.	rep_endemism
# 
en2_taxa_threat <- c(18:21)
# 18.	mam_threat
# 19.	bird_threat
# 20.	amp_threat
# 21.	rep_threat
# 
en2_taxa_small <- c(22:25)
# 22.	mam_small
# 23.	bird_small
# 24.	amp_small
# 25.	rep_small
# 
en3_mb <- c(26, 29, 32, 35)
en3_vp <- c(27, 30, 33, 36)
en3_ae <- c(28, 31, 34, 37)
# 
# 26.	average_mb
# 27.	average_vp
# 28.	average_ae
# 
# 29.	geometric_mb
# 30.	geometric_vp
# 31.	geometric_ae
# 
# 32.	max_mb
# 33.	max_vp
# 34.	max_ae
# 
# 35.	multi_mb
# 36.	multi_vp
# 37.	multi_ae
# 
en4_all <- c(1, 38, 42)
en4_endemism <- c(2, 39, 43)
en4_threat <- c(3, 40, 44)
en4_small <- c(4, 41, 45)
# 
# 38.	vert_all_10
# 39.	vert_endemism_10 
# 40.	vert_threat_10
# 41.	vert_small_10
# 
# 42.	vert_all_110
# 43.	vert_endemism_110
# 44.	vert_threat_110
# 45.	vert_small_110
# 
en5_composites <- c(46:50, 2, 13) # including vert_endemism and plants
# 46.	estes
# 47.	laurance
# 48.	habitats
# 49.	bird_composite
# 50.	damania

en6_norm_all <- c(1, 51)
en6_norm_endemism <- c(2, 52)
en6_norm_threat <- c(3, 53)
en6_norm_small <- c(4, 54)

# 
# 51.	vert_all_sum_norm
# 52.	vert_endemism_sum_norm
# 53.	vert_threat_sum_norm
# 54.	vert_small_sum_norm


```
