---
title: "Appendix S1:  Consequences of under-explored variation in biodiversity indices used for land-use prioritization"
subtitle: "Ecological Applications"
header-includes: \usepackage[nottoc]{tocbibind}
output: 
  bookdown::pdf_document2:
    keep_tex: true
    extra_dependencies: ["float"]
geometry: 
  margin=1in
author: 
  - Christopher L. Crawford^[Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ] $^1$
  
  - Lyndon D. Estes$^2$
  
  - Timothy S. Searchinger$^1$
  
  - David S. Wilcove$^{1, 3}$
date:
  $^1$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ \newline
    $^2$Graduate School of Geography, Clark University, Worcester, MA \newline
    $^3$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ \newline
    \newline  
    \newline
    \today 
editor_options:
  chunk_output_type: console
bibliography: /Users/christophercrawford/Google Drive/Library/library.bib
---
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand\theequation{S\arabic{equation}}

\listoffigures
\listoftables



Testing to see how the pdf compiles. Here's an in text citation: @Estes2016a did this.

# Header One

## Subheader One

### Subsubheader One

Text citing Figure \@ref(fig:nnd-fig).

(ref:caption-nnd-fig) Average distance to nearest neighbor (adjacency) for ensemble mean conversion areas across different categories of variation, assuming 100% weight on biodiversity and constant average yields throughout Zambia (*bp*).

```{r nnd-fig, echo=FALSE, fig.cap = '(ref:caption-nnd-fig)', echo=FALSE, out.height = "80%"}
# \renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
# \renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
plot(cars)
```


See Table \@ref(tab:species-occurrence).

```{r species-occurrence, echo=FALSE}
library(magrittr)
library(kableExtra)
species_occurrence_zambia <- data.frame(
  "Taxa" = c("Mammals", "Birds", "Amphibians", "Reptiles"),
  "All species" = c("252", "738", "94", "226 (GARD)"),
  "Threatened (CR, EN, VU)" = c("11", "20", "1", 
                                "3 (out of 37 assessed 
                                by IUCN in Zambia)"),
  "Small-Ranged (50% smallest)" = c("17", "67", "3", "10 (GARD)"),
  "Small-Ranged OR Threatened" = c("26", "82", "3", "13 (GARD + IUCN)"),
  "Median global range size (km2)" = c("170,600", "471,000", "4,500", "31,500 (GARD)")
)
 

knitr::kable(species_occurrence_zambia, col.names = c("Taxa", "All species", "Threatened \n(CR, EN, VU)", "Small-Ranged \n(50% smallest)", "Small-Ranged \nOR Threatened", "Median global \nrange size (km2)"),
             caption = "Species occurrence in Zambia.", booktabs = TRUE) %>%
  column_spec(2, width = "6em") %>%
  column_spec(3, width = "6em") %>%
  column_spec(4, width = "6em") %>%
  column_spec(5, width = "6em") %>%
  column_spec(6, width = "6em")
```

Citing Equation \@ref(eq:jac-w).

\begin{equation}
J_{w} (a,b) = \frac{\sum_{i=1}^n min(a_{i}, b_{i})}{\sum_{i=1}^n max(a_{i}, b_{i})} (\#eq:jac-w)
\end{equation}




 
